---
title: "DATA 598 Assignment 2"
author: "Shrusti Ghela"
date: "5/4/2022"
output: pdf_document
---
### Forecast the aus_airpassengers dataset in R package fpp3. (If you use this series, use the last 10 years as the test set.)

```{r}
library(fpp3)
head(aus_airpassengers)
```

### 1. Split the series into train and test sets [1 point].

```{r}
air = ts(aus_airpassengers$Passengers, freq=1, start=c(1970))
plot(air, , ylab="aircraft passengers (in millions)", xlab="Year", main="Total annual aircraft passengers for Australia")
```

```{r}
end(air)
```

```{r}
air_train <- window(air, end = c(2006))
air_test <- window(air, start = c(2007))
```

### 2. Identify an ARIMA model for the train set and justify it. At a minimum, consider the model residuals; you may also consider other evidence [1 point]

By simply eyeballing, we see that the data is not stationary
To verify this, check this using KPSS test. 

KPSS test
$H_0$ : Series is stationary
$H_{\alpha}$ :Series is NOT stationary

```{r}
library(tseries)
kpss.test(air_train)
```
p-value = 0.01, p-value < 0.05, $\implies$ we reject the null-hypothesis

```{r}
kpss.test(ts(diff(air_train, differences = 1), freq=1, start=c(1971)))
plot(ts(diff(air_train, differences = 1), freq=1, start=c(1971)))
```
```{r}
kpss.test(ts(diff(air_train, differences = 2), freq=1, start=c(1972)))
plot(ts(diff(air_train, differences = 2), freq=1, start=c(1972)))
```
We will need difference of 2 to convert the time-series to stationary time series. 

```{r}
library(astsa)
acf2(ts(diff(air_train, differences = 2), freq=1, start=c(1972)))
```
One significant lag in ACF suggests a possibility of one MA term and two significant lag in PACF suggest a possibility of two AR terms

```{r}
sarima(air_train, p=2, d=2, q=1)
```
```{r}
sarima(air_train, p=2, d=2, q=0)
```
```{r}
sarima(air_train, p=0, d=2, q=1)
```
Looking at these 3 models, the aic is least for the (0,2,1) model. Looking at the plots as well, we see that for the (0,2,1) model, the Ljung-Box graph suggests higher p-value for all the lags. Compared to the other models, this is the best one. 

Hence, the ARIMA model that I would select would be the (0,2,1) model based on the evidences provided by these models. 

Verifying this result by auto.arima()

```{r}
library(forecast)
auto.arima(air_train)
```
Yayy! I came to the same result!

### 3. Generate a forecast from the model in #1 for the time period of the test set [1 point].
### 4. Plot the train set, the forecast, and the test set [1 point].
```{r}
air_for = sarima.for(air_train, 10, 0, 2, 1)
lines(air_test, type='o')

air_for
```
### 5. Calculate error metrics for the forecast compared to the test set [1 point]. 

```{r}
accuracy(air_for$pred, air_test)
```

